Fish do not need Visas: the transboundary nature of the world’s exploited marine fish species

### 2019-12-25 Manuscript_Draft.Rmd
References update

Distributions of marine species around the world are not bounded by human-made boundaries; rather shaped by contemporary biotic and abiotic factors, as well as their evolutionary history [@Hunch; @NelsonFishesOfTheWorld]. However, management of living marine resources is predicated on the definition of “stocks” that are delineated by human-made spatial boundaries that often do not correspond to biologically-meaningful population units [@Song:2017iua]. The delineation of Economic Exclusive Zones (EEZs) under the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s, for instance, established political boundaries across the distribution of many species, creating shared species between nations. Theories and empirical experience demonstrate fisheries targeting resources that straddle between political boundaries complicate fisheries management and potentially reduce its effectiveness to achieve stated objectives [@REF]. For example, evironmental fluctuation leading to changes in the stock distribution [@Mills] and fishing quotas set by different nations for a shared fishery resource have led to conflict [@Spijkers:2017ij]. Moreover, climate change, which is shifting species distributions, is likely to exacerbate this type of conflict [@Pinsky:2018cb] and presents further challenges for fisheries management [@Mills].

@@

We used four-criteria that combined multiple species distribution data-sources and the world's EEZs to determine if a species can be considered a “shared species”. The first criteria consisted in only considering shared species between neighbouring EEZs, rather than the extended distribution. For the second criteria we considered a species to be present in a grid cell if all data sources predicted positive occurrence. Current species distributions were predicted from occurrence records obtained from publicly available observational databases of the world’s marine species [@REFNEreus] and two structurally distinct species distribution models [@REFGabs; @Zeller] (see Methods). The third criteria determined whether the species were exploited at the grid cell using spatially explicit catch data [@Zeller]. A species was kept if both catch data and predicted species occurrences were positive at the same spatial cells. Finally, the last criteria determined a spatial treshold to confirm that a species occurs in both neighboring EEZs. Once we have identified the shared stocks, we estimated the contribution of the species to fishing revenues and landings using the SAU database [@Zeller] and categorized them according to catch trends [@Kleisner:2013gh].

@@

Overall, we estimated **633** transboundary species in the world, more than double the number previously estimated [@Teh:2015gd], accounting for **65.4**% of the species analyzed (Fig. 1A). Between 2005 and 2014, national fleets targeting shared stocks within EEZs landed an average of **69.2** million tonnes, representing **64**% of the EEZ-global catch, and generating USD **79,745** million in fishing revenue. This is equivalent to **49.8**% of the global revenue of the same period. These values are higher than previously estimated reinforcing the importance of these stocks at the global level [@Teh].

@@

Transboundary species are particulary important for Northern America (mean = 3,901 $\pm$ 4,497 million USD) and Eastern Asia (mean = 2,827 $\pm$ = 2,667 million USD), significantly differing from other regions in terms of revenue (one-way ANOVA; *F*(16,157) = 4.91, *p* < 0.001, $\alpha$ = 0.05) (Fig 1B). The large variation in revenue from these regions is driven by China (USD 6,279 million), in E. Asia and the US (USD 9,605 million) in N. America who contribue 55% and 83% of the region's revenue, respectiveley. Russia (USD 6,361 million), Peru (USD 5,211) and Japan (USD 3,369) complete the top five countries with more revenue from transboundary species. These 5 nations are responsable for 30% of the global revenue from transboundary species. Peru and Russia lead the production of transboundary species harvesting over 5,000 thousand tones (**Suplemental figure Trans_Catch_MAP**). Peru's catches come mainly from Peruvian anchoveta (*Engraulis ringens*), shared with Chile, accounting for 79% of the transboundary species production. Similarly, Russia shares the (currently) largest fishery in the world, Alaskan polllock (*Theragra chalcogramma*) with both Japan and the US [@SOFIA2018]. Despite this, catches are only significantly different in E. Asia (one-way ANOVA; *F*(16,155) = 3.341, *p* < 0.001, $\alpha$ = 0.05), with China, the world's top fish producer with 15.3 million tons landed in 2016 [@SOFIA2018] getting around 1/3 of its capture from transboundary species, followde by Japan (4.2 million tonnes) and South Korea (2.3 million tonnes).

The exploitation category of transboundary species significantly differs from each other (one-way ANOVA *F*(5,446), *f* = 324.5, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05 for all) with most species fished under 50% of their historical maximum peak (Fig.2). This is true for all (but Oceania) continents with no significan difference between them (one-way ANOVA *F*(5,443), *f* = 1.24, *p* = 0.289, $\alpha$ = 0.05), suggesting that the current global exploitation rate of transboundary species could be over the maximum sustainable yield [@Kleisner:2011wn; @PaulyNature2013]. Countries that are today harvesting at this rate (Category C) should be paying extra care in the management of these species (Fig. 2).

@@

While these catch trends do not directly reflect stock status [@Branch:2011br] - e.g. other factors like markets, fishing policies or extreem events also influence catch volume [@PaulyNature2013] - they could ressemble fisheries that are over exploited or collapsed [@Kleisner:2011wn]. Empirical analysis suggest that management of transboundary species will have better outcomes in terms of fish yield and sustainabillity whith cooperation between nations [@REF]. Examples vary between regions and species including Mexico and the United States over Pacific anchovies [@CisnerosMontemayor:2020kv], Norway and Russia over Atlantic cod (*Gadus morhua*) [@Diekert:2010gp], and Namibia and South Africa over hake (*Merluccius spp*) [@Sumaila:2003vw]. Collaboration can also reduce conflict between nations over shared ressources. Transboundary fisheries are abundant and of economic importance for some of the top countries with past fisheries-related conflicts like Canada, the US, China and Russia [@Spijkers:2019cz]. For example, the United States and Canada had a dispute over Pacific salmons (*Oncorhynchus sp*) that lasted over 10 years and was only resolved after these two nations set a series of regulations to compensate losses [@Miller2003]. Since 2007, the European Union, Norway, Iceland, and the Faroe Islands have been at odds over the size and relative allocation of total allowable catches (TACs) for Atlantic mackerel (*Scomber scombrus*) [@Spijkers:2017ij], a species that is transboundary and contributes **698.3** thousand tones to landings and USD **968** million in revenue to the countries involved in the dispute. In addition to past and present conflicts, climate change could exacerbate conflicts over transboundary species as climate change shifts the distribution of marine species around the globe [@Pinsky].

@@

Species can have a cosmopolitan distribution, that is, a broader distribution, or be limited to a specific region, known as endemic [@NelsonFishesOfTheWorld]. Not surprisingly, highly migratory pelagic-oceanic species are the only group that significantly differs in terms of sharing countries (one-way ANOVA *F*(5,109), *f* = 14.67, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) with a median of 40 nations shared per species (Fig. 3).

@@

Due to their large migration patterns and presence in areas beyond national jurisdiction, pelagic-oceanic species such as skipjack tuna (*Katsuwonus pelamis*), Common thresher shark (*Alopias vulpinus*), and tunas (*Thunnus sp.*) are catalogued as stradling species by FAO definitions and globally managed by Regional Fisheries Management Organizations (RFMOs) [@FAO]. Fifteen major RFMOs in the world oversee the management of 99 species, most of them fished in areas beyond national jurisdiction [@CullisSuzuki:2010fi]. We estimate that 81.8% of these species (*n* = 81) are actually shared within neighbouring RFMOs, with the International Commission for the Conservation of Atlantic Tunas (ICCAT) sharing over 40 species with neighbouring RFMOs (**Suplemental figure RFMO_MAP**). The catch from pelagic-oceanic species comes mainly from five South Pacific Islands nations, from which three (Kiribalti, Solomon and Papua New Guinea) are members of the Pacific Nauru Agreement (PNA) (Fig. 3B). The PNA manages 30% of global skipjack tuna [@Aqorau:2018bh], estimated to produce $2 billion worth of fish every year [@https://www.ffa.int/nauru_agreement].

@@

We determine whether or not a species can be considered transboundary. The method is based on a series of criteria to be met and the overlay of species distributions and the spatial boundaries of the world’s EEZs. For the current analysis we work at the species level and adopt the Food and consider transboundary species as those that occur within the Exclusive Economic Zone (EEZ) of neighboring countries [@Munro:2002uf; @FAO]. All data was scaled to 0.5 degree latitude x 0.5 degree longitude grid cells (*n* = ~180,000 grid cells).

@@

We developed a four-criteria method for determining whether or not a species was transboundary. Only species that meet all criteria were considered as transboundary species and the analysis was done only within the EEZs of coastal states. In some cases, criteria also work as indexes to measure the uncertainty in the analysis as they provided continuos index (e.g. from 0 to 1) rather than a discrete, yes or no, index . All of the analysis was done in the statistical software *R version 3.5.2 (2018-12-20)* with the pacages *tidiverse*[@], *data.table*[@], *wesanderson*[@], *janitor*[@], *spdep*[@], *sf*[@], *sp*[@], *rgdal*[@], *tools*[@], *parallel*[@], *rfishbase*[@], *zoo*[@]. All code is available at https://github.com/jepa/FishForVisa.

@@

We define the EEZ boundaries using the SAU shapefile (updated 1 July 2015, available from http://www.seaaroundus.org) that sub-divides EEZs by regions (e.g. Mexico Pacific and Mexico Atlantic) and determine the intersections between polygons using the *sf* *R* package [@SF:package]. When estimating transboundary species, we filter out those shared by the EEZs sub-regions, and when aggregating by country we only counted the species once, if it appeared in more than one sub-region. Species that were present in EEZs that were non-continental territories neighboring other countries were kept (e.g. Argentina and Falkland Islands), but removed when the non-continental territory belonged to the same nation (e.g. Brazil and Fernando de Noronha).

>>>>

Distributions of marine species around the world are not bounded by human-made boundaries; rather shaped by contemporary biotic and abiotic factors, as well as their evolutionary history [@Hutchinson:1957nt; @Nelson:2016kk]. However, management of living marine resources is predicated on the definition of “stocks” that are delineated by human-made spatial boundaries that often do not correspond to biologically-meaningful population units [@Song:2017iua]. The delineation of Economic Exclusive Zones (EEZs) under the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s, for instance, established political boundaries across the distribution of many species, creating shared species between nations. Theories and empirical experience demonstrate fisheries targeting resources that straddle between political boundaries complicate fisheries management and potentially reduce its effectiveness to achieve stated objectives [@REF]. For example, evironmental fluctuation leading to changes in the stock distribution [@Munro:2002uf] and fishing quotas set by different nations for a shared fishery resource have led to conflict [@Spijkers:2017ij]. Moreover, climate change, which is shifting species distributions, is likely to exacerbate this type of conflict [@Pinsky:2018cb] and presents further challenges for fisheries management [@Miller:2013iv].

@@

We used four-criteria that combined multiple species distribution data-sources and the world's EEZs to determine if a species can be considered a “shared species”. The first criteria consisted in only considering shared species between neighbouring EEZs, rather than the extended distribution. For the second criteria we considered a species to be present in a grid cell if all data sources predicted positive occurrence. Current species distributions were predicted from occurrence records obtained from publicly available observational databases of the world’s marine species and two structurally distinct species distribution models [@Asch:2018ca; Reygondeau:2019uh; @Zeller2016] (see Methods). The third criteria determined whether the species were exploited at the grid cell using spatially explicit catch data [@Zeller2016]. A species was kept if both catch data and predicted species occurrences were positive at the same spatial cells. Finally, the last criteria determined a spatial treshold to confirm that a species occurs in both neighboring EEZs. Once we have identified the shared stocks, we estimated the contribution of the species to fishing revenues and landings using the SAU database [@Zeller2016] and categorized them according to catch trends [@Kleisner:2013gh].

@@

Overall, we estimated **633** transboundary species in the world, more than double the number previously estimated [@Teh:2015gd], accounting for **65.4**% of the species analyzed (Fig. 1A). Between 2005 and 2014, national fleets targeting shared stocks within EEZs landed an average of **69.2** million tonnes, representing **64**% of the EEZ-global catch, and generating USD **79,745** million in fishing revenue. This is equivalent to **49.8**% of the global revenue of the same period. These values are higher than previously estimated reinforcing the importance of these stocks at the global level [@Teh:2015gd].

@@

Transboundary species are particulary important for Northern America (mean = 3,901 $\pm$ 4,497 million USD) and Eastern Asia (mean = 2,827 $\pm$ = 2,667 million USD), significantly differing from other regions in terms of revenue (one-way ANOVA; *F*(16,157) = 4.91, *p* < 0.001, $\alpha$ = 0.05) (Fig 1B). The large variation in revenue from these regions is driven by China (USD 6,279 million), in E. Asia and the US (USD 9,605 million) in N. America who contribue 55% and 83% of the region's revenue, respectiveley. Russia (USD 6,361 million), Peru (USD 5,211) and Japan (USD 3,369) complete the top five countries with more revenue from transboundary species. These 5 nations are responsable for 30% of the global revenue from transboundary species. Peru and Russia lead the production of transboundary species harvesting over 5,000 thousand tones (**Suplemental figure Trans_Catch_MAP**). Peru's catches come mainly from Peruvian anchoveta (*Engraulis ringens*), shared with Chile, accounting for 79% of the transboundary species production. Similarly, Russia shares the (currently) largest fishery in the world, Alaskan polllock (*Theragra chalcogramma*) with both Japan and the US [@FAO:2018KK]. Despite this, catches are only significantly different in E. Asia (one-way ANOVA; *F*(16,155) = 3.341, *p* < 0.001, $\alpha$ = 0.05), with China, the world's top fish producer with 15.3 million tons landed in 2016 [@FAO:2018KK] getting around 1/3 of its capture from transboundary species, followde by Japan (4.2 million tonnes) and South Korea (2.3 million tonnes).

The exploitation category of transboundary species significantly differs from each other (one-way ANOVA *F*(5,446), *f* = 324.5, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05 for all) with most species fished under 50% of their historical maximum peak (Fig.2). This is true for all (but Oceania) continents with no significan difference between them (one-way ANOVA *F*(5,443), *f* = 1.24, *p* = 0.289, $\alpha$ = 0.05), suggesting that the current global exploitation rate of transboundary species could be over the maximum sustainable yield [@Kleisner:2011wn; @Pauly:2013gs]. Countries that are today harvesting at this rate (Category C) should be paying extra care in the management of these species (Fig. 2).

@@

While these catch trends do not directly reflect stock status - e.g. other factors like markets, fishing policies or extreem events also influence catch volume [@Pauly:2013gs; @Branch:2008kk] - they could ressemble fisheries that are over exploited or collapsed [@Kleisner:2011wn]. Empirical analysis suggest that in most cases, management of transboundary species will have better outcomes in terms of fish yield and sustainabillity whith cooperation between nations [@Bailey:2010ga]. Examples vary between regions and species including Mexico and the United States over Pacific anchovies [@CisnerosMontemayor:2020kv], Norway and Russia over Atlantic cod (*Gadus morhua*) [@Diekert:2010gp], and Namibia and South Africa over hake (*Merluccius spp*) [@Sumaila:2003vw]. Collaboration can also reduce conflict between nations over shared ressources. Transboundary fisheries are abundant and of economic importance for some of the top countries with past fisheries-related conflicts like Canada, the US, China and Russia [@Spijkers:2019cz]. For example, the United States and Canada had a dispute over Pacific salmons (*Oncorhynchus sp*) that lasted over 10 years and was only resolved after these two nations set a series of regulations to compensate losses [@Miller2003]. Since 2007, the European Union, Norway, Iceland, and the Faroe Islands have been at odds over the size and relative allocation of total allowable catches (TACs) for Atlantic mackerel (*Scomber scombrus*) [@Spijkers:2017ij], a species that is transboundary and contributes **698.3** thousand tones to landings and USD **968** million in revenue to the countries involved in the dispute. In addition to past and present conflicts, climate change could exacerbate conflicts over transboundary species as climate change shifts the distribution of marine species around the globe [@Pinsky:2018cb].

@@

Species can have a cosmopolitan distribution, that is, a broader distribution, or be limited to a specific region, known as endemic [@Nelson:2016kk]. Not surprisingly, highly migratory pelagic-oceanic species are the only group that significantly differs in terms of sharing countries (one-way ANOVA *F*(5,109), *f* = 14.67, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) with a median of 40 nations shared per species (Fig. 3).

@@

Due to their large migration patterns and presence in areas beyond national jurisdiction, pelagic-oceanic species such as skipjack tuna (*Katsuwonus pelamis*), Common thresher shark (*Alopias vulpinus*), and tunas (*Thunnus sp.*) are catalogued as stradling species and globally managed by Regional Fisheries Management Organizations (RFMOs). Fifteen major RFMOs in the world oversee the management of 99 species, most of them fished in areas beyond national jurisdiction [@CullisSuzuki:2010fi]. We estimate that 81.8% of these species (*n* = 81) are actually shared within neighbouring RFMOs, with the International Commission for the Conservation of Atlantic Tunas (ICCAT) sharing over 40 species with neighbouring RFMOs (**Suplemental figure RFMO_MAP**). The catch from pelagic-oceanic species comes mainly from five South Pacific Islands nations, from which three (Kiribalti, Solomon and Papua New Guinea) are members of the Pacific Nauru Agreement (PNA) (Fig. 3B). The PNA manages 30% of global skipjack tuna [@Aqorau:2018bh], estimated to produce $2 billion worth of fish every year [@https://www.ffa.int/nauru_agreement].

@@

We determine whether or not a species can be considered transboundary. The method is based on a series of criteria to be met and the overlay of species distributions and the spatial boundaries of the world’s EEZs. For the current analysis we work at the species level and adopt the Food and consider transboundary species as those that occur within the Exclusive Economic Zone (EEZ) of neighboring countries [@Munro:2002uf]. All data was scaled to 0.5 degree latitude x 0.5 degree longitude grid cells (*n* = ~180,000 grid cells).

@@

We developed a four-criteria method for determining whether or not a species was transboundary. Only species that meet all criteria were considered as transboundary species and the analysis was done only within the EEZs of coastal states. In some cases, criteria also work as indexes to measure the uncertainty in the analysis as they provided continuos index (e.g. from 0 to 1) rather than a discrete, yes or no, index . All of the analysis was done in the statistical software *R version 3.5.2 (2018-12-20)* with the pacages *tidiverse*[@PackagetidyverseE:2017vq], *data.table*[@Packagedatatable:2019uh], *wesanderson*[@], *janitor*[@], *spdep*[@], *sf*[@PackagesfSimpleF:2018vp], *sp*[@], *rgdal*[@PackagergadalBind:2018vy], *tools*[@], *parallel*[@], *rfishbase*[@], *zoo*[@]. All code is available at https://github.com/jepa/FishForVisa.

@@

We define the EEZ boundaries using the SAU shapefile (updated 1 July 2015, available from http://www.seaaroundus.org) that sub-divides EEZs by regions (e.g. Mexico Pacific and Mexico Atlantic) and determine the intersections between polygons using the *sf*-*R* package. When estimating transboundary species, we filter out those shared by the EEZs sub-regions, and when aggregating by country we only counted the species once, if it appeared in more than one sub-region. Species that were present in EEZs that were non-continental territories neighboring other countries were kept (e.g. Argentina and Falkland Islands), but removed when the non-continental territory belonged to the same nation (e.g. Brazil and Fernando de Noronha).
